Paper 1 Analysis

### 2015-10-22 Articles/Article_EcolEcon_2015/EcolEcon_2015A.Rnw
Now making better graphs.

print("top of Fitted_and_Residuals")

temp_data_exp <- 
  leaf_apply(
    modelsExp, class=c("lm"), 
    f = function(model, id){
      model$dat %>% 
        select(Year, Country) %>%
        mutate(
          Country = as.character(Country),  # to avoid warning about unmatched levels of factors
          fitted = exp(fitted(model)), # exp accounts for the fact that the model was fitted in log space.
          residuals = resid(model)     
        )
    }
  ) %>% 
  bind_rows()

### If the function is this long, it should be separated out and named rather than embedded in another
### function call.  
### But this looks like model2df from above.  Let's use that instead.

### Probably just junk now.
# function_that_was_here_before <- function(model, id){
#   df <- data.frame(select(model$data, Year, Country, iGDP))
#   df$fitted <- exp(fitted(model)) # fitted is in log space. Return to regular space.
#   df$Residuals <- resid(model) # residuals are in log space, and that is what we want.
#   # Now add meta information
#   id_list <- strsplit(id, split = "[.]")[[1]]
#   df$flavor <- if (id_list[[2]] ==  "unadjusted") "Unadjusted" else "Quality-adjusted"
#   df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
#   df$model <- id_list[[4]]
#   df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
#   return(df)
# }

### This looks like a candidate for clean up a la model2df

@@

temp_data_noExp <- 
  bind_rows( temp_data_CSP, temp_data_noCSP )

# Now merge the exponential-only model with the CSP and noCSP models. 
# Doing this, because the exponential-only model is kind-of like iY:
# it applies to all situations.
fitted_and_resid_data <- merge(temp_data_noExp, temp_data_exp)

# Relevel factors before plotting
fitted_and_resid_data <-
  fitted_and_resid_data %>% 
  mutate(
    Country = relevelFactor(factor(Country), c("PT", "UK")),
    flavor  = relevelFactor(factor(flavor), c("Unadjusted", "Quality-adjusted")),
    energy  = relevelFactor(factor(energy), c("Without energy", "With energy")),
    nest    = relevelFactor(factor(nest), c("kl", "kle", "lek", "ekl"))
  ) %>%
  arrange(Country, CSP, model, flavor, energy, nest)
@

@@

<<Fitted_and_Residual_Graph_Functions>>=

### Can fittedGraphNoCSP and fittedGraphCSP be combined into one function with an argument?
### Or can the data filtering happen outside and everything else be the same?
### Did you copy and paste to make one from the other?  That should be a sign...
fittedGraphNoCSP <- function(data){
  fittedGraphData <- 
    data %>% 
    select(Year, Country, iGDP, fitted, fittedExp, flavor, energy, CSP, nest) %>%
    melt(measure.vars = c("iGDP", "fitted", "fittedExp"), 
         value.name = "value", variable.name = "variable") %>%
    mutate(label = paste(variable, energy)) %>%
    # Now collapse some of the information. Don't need multiple versions of the exponential model, e.g.
    filter(label %in% c("iGDP With energy", "fittedExp With energy", 
                        "fitted With energy", "fitted Without energy")
    ) %>% 
    mutate(
      label = plyr::mapvalues(label, 
                              from = c("iGDP With energy",  "fittedExp With energy",
                                       "fitted With energy", "fitted Without energy"), 
                              to = c("GDP",  "Ref. model",
                                     "w/ energy", "w/o energy")),
      label = relevelFactor(factor(label), 
                            c("GDP", "Ref. model", "w/o energy", "w/ energy"))
    )

  ggplot() +
    geom_line(data=fittedGraphData, mapping=aes(x=Year, y=value, group=label, linetype=label, size=label)) +
    scale_linetype_manual(values=c(1,2,4,3), name="") +
    scale_size_manual(name = "", values=c(1, 0.5, 0.5, 0.5)) +
    facet_grid(flavor ~ Country) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

<<ResidualPlotFunction>>=

cesResidualPlot2 <- function(data){
  dataToPlot <- 
    data %>%
    mutate(nest = 
             plyr::mapvalues(
               nest, 
               from=c("kl", "kle", "lek", "ekl"), 
               to=c("kl", "(kl)e", "(le)k", "(ek)l")
               ),
           Model = interaction(nest, CSP, sep = " ", drop=TRUE),
           Inputs = flavor)

  ggplot() +
    geom_point(
      data=dataToPlot, 
      mapping=aes(
        x=Year, 
        y=Residuals, 
        fill=Model, 
        shape=Model, 
        size=Model
        )
      ) +
    facet_grid(flavor ~ Country) +
    geom_hline(yintercept = 0, colour="gray50", size=0.1) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xlab("") +
    ylab("r") + 
    xy_theme() +
    labs(color = "Factors of production") +
    # This is custom stuff here to match the model_points_style function. 
    # We can't use model_points_style, because the aesthetics are different here
    # with only two series (kl w/o CSP and (kl)e w/o CSP) in the graph.
    scale_size_manual(values = c(3, 3)) +
    scale_fill_manual(values = c("transparent", "transparent")) +
    scale_shape_manual(values = c(21, 23)) +
  theme(legend.title=element_blank())
}
@

>>>>

temp_data_exp <- 
  leaf_apply(
    modelsExp, class=c("lm"), 
    f = function(model, id){
      out <- model$dat %>% 
        select(Year, Country) %>%
        mutate(
          Country = as.character(Country),  # to avoid warning about unmatched levels of factors
          fitted = exp(fitted(model)), # exp accounts for the fact that the model was fitted in log space.
          residuals = resid(model), 
          model = "exp"
        )
        # Now add the response, using the original variable name.
      out[[summands(model$formula, right = FALSE)]] <- exp(response(model)[c(1:length(out$Year))])
      return(out)
    }
  ) %>% 
  bind_rows()

@@

FittedAndResid <- plyr::rbind.fill(temp_data_exp, temp_data_CSP, temp_data_noCSP)

@@

### 2015-10-22 Articles/Article_EcolEcon_2015/Paper1/Paper1-lambda-alpha.Rnw
Now making better graphs.

<<Graphs_CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and $(kl)e$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_kle <- 
  fitted_and_resid_data %>% 
  filter(model=="CES" & CSP=="w/o CSP" & (nest == "kl" | nest == "kle"))

fittedGraphNoCSP(fittedGraphData_kle)
@

<<Fitted_Graph, eval=FALSE>>=
fitted_and_resid_data %>% 
  filter(model=="CES") %>%
  gather()
  melt(measure.vars = c("iGDP", "fitted", "fittedExp"), 
       value.name = "value", variable.name = "variable") %>%
  mutate(label = paste(variable, energy)) %>%
  # Now collapse some of the information. Don't need multiple versions of the exponential model, e.g.
  filter(label %in% c("iGDP With energy", "fittedExp With energy", 
                      "fitted With energy", "fitted Without energy")
  ) %>% 
  mutate(
    label = plyr::mapvalues(label, 
                            from = c("iGDP With energy",  "fittedExp With energy",
                                     "fitted With energy", "fitted Without energy"), 
                            to = c("GDP",  "Ref. model",
                                   "w/ energy", "w/o energy")),
    label = relevelFactor(factor(label), 
                          c("GDP", "Ref. model", "w/o energy", "w/ energy"))
  ) %>%

  ggplot() +
  geom_line( mapping=aes(x=Year, y=value, group=label, linetype=label, size=label)) +
  scale_linetype_manual(values=c(1,2,4,3), name="") +
  scale_size_manual(name = "", values=c(1, 0.5, 0.5, 0.5)) +
  facet_grid(flavor ~ Country) +
  scale_x_continuous(breaks=c(1960, 1980, 2000)) +
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  xlab("") +
  ylab("Indexed GDP (1960=1)") +
  xy_theme()

@@

<<SRsse_CES_graph_3, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Solow Residuals~($\\lambda$) and mean squared error~($sse$). $\\lambda$ and $sse$ values for the reference (exponential-only) model are shown as gridlines. This version uses our proposed new style.">>=
data_sr_sse <- 
  AllCoeffs %>% 
  filter(model=="CES") %>% 
  mutate(nest = 
           plyr::mapvalues(
             nest, 
             from=c("kl", "kle", "lek", "ekl"), 
             to=c("kl", "(kl)e", "(le)k", "(ek)l")
             ),
         Model = interaction(nest, CSP, sep = " ", drop=TRUE),
         Inputs = flavor
  )

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fitted Graphs}
\label{sec:appendix_fitted}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<Graphs_CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and $(le)k$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_lek <- 
  fitted_and_resid_data %>% 
  filter(model=="CES" & CSP=="w/o CSP" & (nest == "kl" | nest == "lek"))

fittedGraphNoCSP(fittedGraphData_lek)
@

<<Graphs_CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and $(ek)l$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_ekl <- 
  fitted_and_resid_data %>% 
  filter(model=="CES" & CSP=="w/o CSP" & (nest == "kl" | nest == "ekl"))

fittedGraphNoCSP(fittedGraphData_ekl)
@

@@

<<Graphs_CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(kl$)$e$ nesting.">>=
cesResidualPlot2(
  fitted_and_resid_data %>% 
    filter(model=="CES" & CSP=="w/o CSP" & (nest=="kle" | nest=="kl"))
)
@

<<Graphs_CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with ($le$)$k$ nesting.">>=
cesResidualPlot2(
  fitted_and_resid_data %>% 
    filter(model=="CES" & CSP=="w/o CSP" & (nest == "lek" | nest=="kl"))
)
@

<<Graphs_CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(ek$)$l$ nesting.">>=
cesResidualPlot2(
  fitted_and_resid_data %>% 
    filter(model=="CES" & CSP=="w/o CSP" & (nest=="ekl" | nest=="kl"))
)
@

>>>>

<<faceting_functions>>=

# Replicates exponential models for each flavor in the flavors vector
replicateExpModels <- function(data, flavors){
  lapply(flavors, function(flav){
    data %>% 
      filter(model=="exp") %>%
      mutate(
        flavor = flav,
        nest = "exp",
        CSP = "exp"
      )
  }) %>%
    bind_rows() %>%

    # Combine the replicated exponential models with everything else in data
    rbind(., filter(data, model != "exp"))
}

# Add facet columns to data for both fitted and residuals graphs.
addFacetCols <- function(data){
  # Create the facet variables
  mutate(data,
    nest = plyr::mapvalues(nest, from=c("kle", "lek", "ekl"), to=c("(kl)e", "(le)k", "(ek)l")),
    flavor = plyr::mapvalues(flavor, from=c("Unadjusted", "Quality-adjusted"), to=c("UA", "QA")),

    xfacet = interaction(Country, flavor, sep = " - "),
    xfacet = relevelFactor(factor(xfacet), c("PT - UA", "PT - QA", "UK - UA", "UK - QA")),

    yfacet = interaction(model, nest, CSP, sep = " - "),
    yfacet = plyr::mapvalues(yfacet, 
                             from=c("exp - exp - exp", 
                                    "CES - kl - w/ CSP",
                                    "CES - kl - w/o CSP",
                                    "CES - (kl)e - w/o CSP",
                                    "CES - (le)k - w/o CSP",
                                    "CES - (ek)l - w/o CSP"), 
                             to  =c("Ref.",
                                    "kl - w/ CSP",
                                    "kl - w/o CSP",
                                    "(kl)e - w/o CSP",
                                    "(le)k - w/o CSP",
                                    "(ek)l - w/o CSP")
    ),
    yfacet = relevelFactor(factor(yfacet), c("Ref.", 
                                             "kl - w/ CSP", 
                                             "kl - w/o CSP", 
                                             "(kl)e - w/o CSP", 
                                             "(le)k - w/o CSP", 
                                             "(ek)l - w/o CSP"))
  )
}

@@

<<Fitted_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=7, fig.cap="Fitted compared to historical GDP time series.">>=

# Duplicate the exponential models for both Unadjusted and Quality-adjusted cases.
replicateExpModels(
  data = filter(FittedAndResid, model=="exp" | model=="CES"), 
  flavors = c("Unadjusted", "Quality-adjusted")
  ) %>%

  addFacetCols() %>%

  gather(key = y_type, value = y_value, iGDP, fitted) %>%

  ggplot() +
  geom_line(alpha = 0.8, aes(x = Year, 
                             y = y_value,
                             group = y_type,
                             color = y_type,
                             linetype = y_type,
                             size = y_type)
  ) + 
  facet_grid(yfacet ~ xfacet) + 
  scale_x_continuous(breaks = c(1960, 1980, 2000)) +
  scale_color_manual(
    name = "",
    labels = c("historical", "fitted"),
    values=c("gray70", "black")
  ) +
  scale_size_manual(
    name = "",
    labels = c("historical", "fitted"),
    values = c(1.5, 0.3)) +
  scale_linetype_manual( 
    name = "",
    labels = c("historical", "fitted"),
    values=c(1,1) ) + 
  labs(x="", y="GDP") +
  xy_theme()

<<SRsse_CES_graph_3, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Solow Residuals~($\\lambda$) and mean squared error~($sse$). $\\lambda$ and $sse$ values for the reference (exponential-only) model are shown as gridlines. This version uses our proposed new style.">>=
data_sr_sse <- 
  AllCoeffs %>% 
  filter(model=="CES") %>% 
  mutate(nest = 
           plyr::mapvalues(
             nest, 
             from=c("kl", "kle",   "lek",   "ekl"), 
             to  =c("kl", "(kl)e", "(le)k", "(ek)l")
             ),
         Model = interaction(nest, CSP, sep = " ", drop=TRUE),
         Inputs = flavor
  )

@@

@@

<<Residuals_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=7, fig.cap="Model residuals.">>=

# Duplicate the exponential models for both Unadjusted and Quality-adjusted cases.
replicateExpModels(
  data = filter(FittedAndResid, model=="exp" | model=="CES"), 
  flavors = c("Unadjusted", "Quality-adjusted")
  ) %>%

  addFacetCols() %>%

  ggplot() +
  geom_point(alpha = 0.8, size = 0.5, aes(x = Year, y = residuals)) + 
  facet_grid(yfacet ~ xfacet, scales = "free_y" ) + 
  scale_x_continuous(breaks = c(1960, 1980, 2000)) +
  labs(x="", y="Residuals") +
  geom_hline(aes(yintercept = 0.0), colour="gray50", size=0.1) + 
  xy_theme()
